{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation

### 2021-03-14 documents/summary.tex
Screw it, let's try liposomes

%\begin{refsection}
\lettrine{W}e were inspired to undertake this project by a set of publications by a certain Sun group, (primarily Liu et al \cite{Microwave2009} and Yang et al \cite{Efficient2015}, simultaneously demonstrated by Hung et al \cite{focusing2014}), partially replicated by \cite{Optical2020} and discussed by \cite{Theoretical2020}.

@@

A few well-established non-resonant non-thermal effects caused by extreme electric fields have already been put to practical use. Electroporation\cite{Electroporation1988}, now a common laboratory procedure, occurs when an enormous electric field causes ions in solution to diffuse across a cell membrane, capacitively charging it. This leads to a change in lipid conformation, reversibly rendering the membrane permeable. This is one among other, more subtle mechanisms; the minutia of poration are extraordinarily complex and beyond the author's current understanding\cite{Theoretical2007}.

Therapies based on irreversible electroporation\cite{Nonthermal2013}\cite{Lipid2017} have recently been used in the clinic\cite{Irreversible2013} for tumor ablation. There is much discussion on electroporation as an alternative to viral vectors, but we have found surprisingly little discussion on clinical treatment of viral infection with these conventional electroporation therapies.% In vitro there is evidence to suggest that sub-cellular effects take place under truly tremendous fields.

Long pulses at $\approx10$ MV/m appear to electroporate enveloped viruses similarly to host cells\cite{AC2017}, which we believe might hint that a regime of intensity, duration, and pulse count may exist to target the viral envelope. The same field intensity does not appear to damage phages with protein capsids\cite{Manipulation2013}.

@@

However, perhaps regrettably, the experiments reported here were not designed to address the question of whether conventional ion- and bilayer- driven electroporation can be harnessed, but rather to test whether purely mechanical damage from $qE$ forces can be easily observed.

@@

Almost all biological spectroscopy techniques have produced resonance-like artifacts at some point in the literature, and we suggest that even the most striking and competently-obtained absorption data should not necessarily be trusted implicitly. Trustworthiness probably depends most significantly on the quality of error analysis; many have made good use of otherwise unreliable techniques. However, particularly in a field with a high positive publication bias, it seems most productive to use the most advanced techniques, those that are foolproof by their very nature. Based only on our experience of the literature, in order of decreasing reliability for resonance measurement:

\begin{itemize}
\item Background-free photothermal optical \cite{Microwave1993a}\cite{Broadband1988}
\item Time-domain spectroscopy (many advantages cited by \cite{Time2003}, including instant, drift-free capture of all frequencies and limited instrumentation artifacts), \cite{Dielectric2004}\cite{Microwave1994}
\item Raman or other non-photothermal optical (\cite{optical1983} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) and also \footnote{see Taylor \cite{mechanisms1981} citing unpublished, personal communication}, but we have no understanding of how such artifacts could form
\item Frequency-domain CW VNA absorption or reflection (\cite{Microwave1982} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) \cite{Substitution1982} \cite{Millimeter1980}.
\end{itemize}

Although not representative of careful, well-calibrated VNA testing because of the crudity of our homebrew spectrometer, we were able to produce many candidate resonances in our data with perfectly justifiable modifications to post-processing, which raises interesting (but not novel\cite{Reanalysis2020}\cite{statistical2020}) questions regarding software development in such conditions. \footnote{One solution might be to keep all data blinded while developing the analysis software, and use synthetic data to calibrate. One objection might be that the data often inform the required calibration; we might retort that if the technique is so sensitive that the calibration procedure cannot be known beforehand, a different technique should be used.}

@@

\subsection{Materials and methods}

@@

Positive control samples were put in a commercial pressure cooker to autoclave for either 6 or 20 minutes in an Eppendorf tube: the pressure was released after the cycle to prevent dilution. The flurophore was added after autoclaving.

@@

Avalanche transistors appear to be the easiest way to produce high power sub-100 picosecond pulses. We stole an avalanche Marx design verbatim from Li (\cite{Development2016b} \cite{Design2018c}), which allows large pulses from safer lower voltage supplies (albeit requiring somewhat costly capacitors). 2N5551 transistors used to be common \cite{Avalanche1991}\cite{high1994}\cite{High1998}; our testing with Raj's FMMT417

@@

A programmable syringe withdrew and pumped 50 uL of fluid at a constant rate through one of the coplanar channels, via a 4 cm long silicone capillary (McMaster-Carr \#51845K65) immersed in a 1.5 mL centrifuge tube using a 26 Ga Luer-lock (\#6710A76). Forcing phage through the capillary is similar to lysis with a French pressure cell press, but no such effects were obviously seen.

@@

In all cases, a thermal effect was made implausible; the total energy in the pulse train precluded heating above the capsid melting temperature\cite{Effects1951}.

@@

We are particularly excited by the prospect of using picosecond time-domain measurements not only to perform steady state dielectric measurements or monitor the input pulse, as is now common practice, but to resolve the internal effects of a single high-voltage pulse on the virus in real time. Scan-converter based transient digitizers, such as the Tektronix SCD-5000 series, appear to be the lowest-cost method of obtaining the >0.5 THz sampling bandwidths required, when the tube bandwidth is deconvolved.

@@

Funded in part by the author's company, SafeSump Inc., and Ideal Consulting Ltd., and the author's parents. The author was at YorkU for the duration of this project.

@@

\subfile{brillouin_pulse_math}

>>>>

%\begin{refsection}

A few well-established non-resonant non-thermal phenomena caused by extreme electric fields have been put to practical use.

Electroporation\cite{Electroporation1988}, now a common laboratory procedure, occurs when an enormous electric field causes ions in solution to diffuse across a cell membrane, capacitively charging it, provoking a change in lipid bilayer conformation\cite{Membrane2016}, reversibly rendering the membrane permeable. This is one among many other, more subtle mechanisms; the minutia of poration can become extraordinarily complex in certain circumstances and are beyond the author's current understanding\cite{Theoretical2007}.

Therapies based on irreversible electroporation\cite{Nonthermal2013}\cite{Lipid2017} have recently been used in the clinic\cite{Irreversible2013} for tumor ablation. There is much discussion on electroporation as an alternative to viral vectors, but we have found surprisingly little discussion on clinical treatment of viral infection with these conventional electroporation therapies.

However, long pulses at $\approx10$ MV/m appear to electroporate large enveloped viruses, similarly to host cells\cite{AC2017}, which we believe might hint that a regime of intensity, duration, and pulse count may exist to target the viral envelope. In fact, despite considering the somewhat harder problem of electroporating while inside the cell, rather than in extracellular fluid, \cite{Electroporation2013} report selective electroporation of 250 nm liposomes with 4 ns pulses up to 10 MV/m without inducing damage to the host cell - which seems a very important result. Despite the number of free parameters in pulse exposure, it is possible that scaling laws might conceivably preclude the formation of destructive transmembrane potentials in 100 nm-scale viruses.

The same field intensities do not appear to damage phages with protein capsids\cite{Manipulation2013}, in line with our negative results on T4.

However, perhaps regrettably, the experiments reported in this study were not designed to address the question of whether conventional ion- and bilayer- driven electroporation can be harnessed, but rather to test whether purely mechanical damage from $qE$ forces can be easily observed in viruses.

\lettrine{W}e were inspired to undertake this project by a set of publications by a certain Sun group, (primarily Liu et al \cite{Microwave2009} and Yang et al \cite{Efficient2015}, simultaneously demonstrated by Hung et al \cite{focusing2014}), partially replicated by \cite{Optical2020} and discussed by \cite{Theoretical2020}.

@@

@@

@@

Almost all biological spectroscopy techniques have produced resonance-like artifacts at some point in the literature, and we suggest that even the most striking and competently-obtained absorption data should not necessarily be trusted implicitly. Trustworthiness probably depends most significantly on the quality of error analysis; many have made good use of otherwise unreliable techniques. However, particularly in this field, it seems most productive to use the most advanced techniques, those that are foolproof by their very nature. Based only on our experience of the literature, in order of decreasing reliability for resonance measurement:

\begin{itemize}
\item Background-free photothermal optical \cite{Microwave1993a}\cite{Broadband1988}
\item Time-domain spectroscopy (many advantages cited by \cite{Time2003}, including instant, drift-free capture of all frequencies and limited instrumentation artifacts), \cite{Dielectric2004}\cite{Microwave1994}
\item Raman or other non-photothermal optical (\cite{optical1983} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) and also \footnote{see Taylor \cite{mechanisms1981} citing unpublished, personal communication - a textbook example of positive publication bias!}, but we have no understanding of how such artifacts could form
\item Frequency-domain CW VNA absorption or reflection (\cite{Microwave1982} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) \cite{Substitution1982} \cite{Millimeter1980}.
\end{itemize}

Although not representative of careful, well-calibrated VNA testing because of the crudity of our homebrew spectrometer, we were able to produce many candidate resonances in our data with perfectly justifiable modifications to post-processing, which raises interesting (but not novel\cite{Blindinga}, \cite{Reanalysis2020}\cite{statistical2020}) questions regarding software development in such conditions. \footnote{One solution might be to keep all data blind while developing the analysis software, and use synthetic data to calibrate. One objection might be that the data often inform the required calibration; we might retort that if the technique is so sensitive that the calibration procedure cannot be known beforehand, a different method with a higher SNR should be used.}

@@

\subsection{Experiment notes}

@@

Positive control samples were put in a commercial pressure cooker to autoclave for 20 minutes in an Eppendorf tube: the pressure was released after the cycle to prevent dilution. The flurophore was added after autoclaving.

@@

Avalanche transistors appear to be the easiest way to produce high power sub-100 picosecond pulses. We stole an avalanche Marx design verbatim from Li et al (\cite{Development2016b} \cite{Design2018c}), which allows large, fast pulses to be produced from lower voltage supplies (albeit requiring somewhat costly capacitors). 2N5551 transistors used to be commonly used for this purpose. \cite{Avalanche1991}\cite{high1994}\cite{High1998}; Oak's test circuit .

@@

A programmable syringe withdrew and pumped approximately 50 uL of fluid at a constant rate through one of the coplanar channels, via a 4 cm long silicone capillary (McMaster-Carr \#51845K65) immersed in a 1.5 mL centrifuge tube using a 26 Ga Luer-lock (\#6710A76). Forcing phage through the capillary is similar to lysis with a French pressure cell press, but no such effects were obviously seen.

@@

In all cases, a thermal effect was made implausible because of the short pulse duration; the total energy in the pulse train precluded heating above the capsid melting temperature, like\cite{Effects1951} - about 90 C for capsids \cite{Thermal1999} and 50 C for envelopes\cite{Stability1985}.

@@

We are particularly excited by the prospect of using picosecond time-domain measurements not only to perform low-field-intensity steady state dielectric measurements or monitor the input pulse, as is now common practice, but to resolve the internal effects of a single high-voltage pulse on the virus in real time. Scan-converter based transient digitizers, such as the Tektronix SCD-5000 series, appear to be the lowest-cost method of obtaining the >0.5 THz sampling bandwidths required, when the tube bandwidth is deconvolved.

Fecher\cite{Production1949} micro-oscillographs

\section{Materials and methods}

\subsubsection{Main sample preparation}

A stock preparation of 1:150 Biotium GelGreen was diluted to 1:4000 in a 15 mL Falcon tube.

With a Kartell 0.2 mL fixed-volume micropipette (McMaster-Carr \#7015T17) connected to an autoclaved and flamed glass pipette,  0.2 mL of Carolina 124330 (early assay tests) or Carolina 124335 (electric field tests), titer $3\times10^9$, was transferred into a 1.5 mL Eppendorf tube. Field-tested phage was stored in a 4 C fridge for several months during the experiment.

 stored in an aluminum light-tight bag was pipetted into the same 1.5 mL tube.

\subsubsection{Main sample preparation}

\subsubsection{Nanosecond pulsed trials}

A clone of Li et al's pulse generator was engraved with a 30 degree V-bit cutter onto a 1.6 mm double-sided 1 oz copper FR4 substrate. ATC 100B470JT500XT capacitors and FMMT417 transistors were used.

Two coplanar waveguides, approximately 2.5 cm long, designed with 'wcalc'\cite{Wcalc}, were similarly engraved into a single-sided 0.8 mm 1 oz copper FR4 substrate (MG chemicals). One waveguide (referred to as "0.2 mm channel") was 0.4 mm wide with 0.1 mm and 0.15 mm wide slots (asymmetric due to machining tolerances). The other was 0.4 mm wide with 0.03 and 0.06 mm slots.

The depth of the channels was not precisely known but considerably exceeded the copper thickness, as is reflected in the simulation.

Two end-launch SMA connectors were soldered to the waveguide; one connector was either terminated with a standard 50-ohm SMA terminator or left open, depending on the trial, and the other connected via semirigid coax (CCSMA-MM-086-12) to the pulse generator.

\conclusiongrad

HIA4 power supply.

\subsubsection{FDTD tissue modelling}

To determine penetration of waves in tissue, the four term Cole-Cole model\cite{gabriel1996compilation}, \ghfile{/electronics/simple_fdtd/fdtd_PCB_extensions/fdtd_PCB_extensions/tissue.py}, was used with parameters from the IT.IS Foundation tissue database\cite{Tissue2018a}. The IT.IS Foundation computational phantom 'duke V3.0' was used, \cite{Duke2014}.

@@

Funded in part by the author's company, SafeSump Inc., and Ideal Consulting Ltd., and the author's parents.

@@

%\subfile{brillouin_pulse_math}
